Adapting Duplicate Filters to a Sliding Window Context

### 2020-03-20 qhtv2.tex

\subsection{FNR and FPR}
As stated before, the probabilty of false positive and false negative are not the usual measurement made when benchmarking filters. Rather, one prefers to use false positive (and negative) \emph{rates}, defined as the average of the false positive (and negative) probabilities encountered so far.

In this section, we briefly give the expression of these values. For the sake of simplicity, we only give the FPR and FNR of a queueing filter $\mathcal Q$ after $m = cn$ insertions, where $n$ is a positive integer.
\begin{theorem}
	Let $\FPR_{\mathcal Q, m}^w$ be the false positive rate of $\mathcal Q$ after $m=cn > w$ insertions on a sliding window of size $w \approx cL$, we have
	$$\FPR_{\mathcal Q, cn}^w = 1 -  (1 - \FP_{\mathcal F, c})^{L-1}\sum_{\ell = 0}^{c-1} (1-\FP_{\mathcal F, \ell})$$
\end{theorem}
\begin{proof}
\begin{align*}
	\FPR_{\mathcal Q, cn}^w
	& = \frac{1}{cn} \sum_{k=1}^{cn} \FP_{\mathcal Q, k}^w 
	 = \frac{1}{cn} \sum_{k=1}^{n} \sum_{\ell = 0}^{c-1} \FP_{\mathcal Q, k + \ell}^w 
	 = \frac1c \sum_{\ell = 0}^{c-1} \FP_{\mathcal Q, \ell}^w \\
	& = \frac1c \sum_{\ell = 0}^{c-1}  1 - (1 - \FP_{\mathcal F, c})^{L-1}(1 - \FP_{\mathcal F}, \ell) \\
	& = 1 -  (1 - \FP_{\mathcal F, c})^{L-1}\sum_{\ell = 0}^{c-1} (1-\FP_{\mathcal F, \ell})
\end{align*}
\end{proof}

@@

	with the $u_\eta= p_{\eta}\FN_{\mathcal F,\eta} + \left(1-p_{\eta}\right)\left(1-\FP_{\mathcal F,\eta}\right)$.
\end{theorem}
\begin{proof}
\begin{align*}
	\FNR_{\mathcal Q, cn}^w 
	& = \frac{1}{cn} \sum_{k=1}^{cn} \FN_{\mathcal Q, k}^w 
	 = \frac{1}{cn} \sum_{k=1}^{n} \sum_{\ell = 0}^{c-1} \FN_{\mathcal Q, k + \ell}^w 
	= \frac1c \sum_{\ell = 0}^{c-1} \FN_{\mathcal Q, \ell}^w \\
	& = \frac1c \sum_{\ell = 0}^{c-1} u_c^{L-1} u_{\ell} 
	 = \frac{u_c^{L-1}}{c}  \sum_{\ell = 0}^{c-1}u_\ell
\end{align*}
\end{proof}

\section{Optimising queuing filters}

\begin{theorem}
	In first approximation, a low value of $L$ for constant value $cL$ leads to a higher false negative probability for the queueing filter.
\end{theorem}

@@

	On the other hand, a duplicate element $e^\star \dup E$ is classified as \UNSEEN if and only if it was not seen in the last $M$ elements of the stream, and the filter answers \UNSEEN. Let $D$ be the event "\textsf{There is at least one duplicate in the stream} and $C$ be the event \textsf{There is a duplicate of $e^\star$ in the $M$ previous elements of the stream}.

	We have that $e^\star$ triggers a false negative with probability

	\begin{align*}
	\FN_n &= (1 - P[C | D])(1-p) =  \left(1 - \frac{P[C \cap D]}{P[D]}\right)(1-p)\\
	& = \left(1 - \frac{P[C]}{P[D]}\right)(1-p) =  \left(1 - \frac{1 - P[\bar C]}{1 - P[\bar D]}\right)(1-p) \\
	\FN_n &=  \left(1 - \frac{1 - \left(1 - \frac 1{|\Gamma|}\right)^M}{1 - \left(1 - \frac 1{|\Gamma|}\right)^n}\right)(1-p)
	\end{align*}

	Hence, the error probability of the filter is $ER_n = \FN_n + p =  \left(1 - \frac{1 - \left(1 - \frac{1}{|\Gamma|}\right)^{M}}{1 - \left(1 - \frac 1{|\Gamma|}\right)^n}\right)(1-p) + p =$ $ 1 - \frac{1-\left(1 - \frac{1}{|\Gamma|}\right)^{M}}{1 - \left(1 - \frac 1{|\Gamma|}\right)^n}(1-p)$, which is minimized when $p = 0$.

	Given that, by definition, a perfect filter has the lowest error rate of any filter, we get the desired result.
	\end{proof}

Note, as highlighted in the proof, that this bound is not tight, and that better bounds may exist. Also note that this bound only applies when the stream is random: when patterns are found in the stream, information theory informs that data may be stored more efficiently.

As such, the best filter will be the one the closest to this error rate.

@@

>>>>

\subsection{FNR and FPR}

As stated before, the probabilty of false positive and false negative are not the usual measurement made when benchmarking filters. Rather, one prefers to use false positive (and negative) \emph{rates}, defined as the average of the false positive (and negative) probabilities encountered so far.

From the above expressions we can derive relatively compact explicit formulas for the queuing filter's FPR and FNR. For the sake of simplicity, we only give the FPR and FNR of a queueing filter $\mathcal Q$ after $m = cn$ insertions, where $n$ is a positive integer.
\begin{theorem}
	Let $\FPR_{\mathcal Q, m}^w$ be the false positive rate of $\mathcal Q$ after $m=cn > w$ insertions on a sliding window of size $w \approx cL$, we have
	$$\FPR_{\mathcal Q, cn}^w = 1 -  (1 - \FP_{\mathcal F, c})^{L-1}\sum_{\ell = 0}^{c-1} (1-\FP_{\mathcal F, \ell})$$
\end{theorem}
\begin{proof}

\begin{align*}
	\FPR_{\mathcal Q, cn}^w
	& = \frac{1}{cn} \sum_{k=1}^{cn} \FP_{\mathcal Q, k}^w 
	 = \frac{1}{cn} \sum_{k=1}^{n} \sum_{\ell = 0}^{c-1} \FP_{\mathcal Q, k + \ell}^w 
	 = \frac1c \sum_{\ell = 0}^{c-1} \FP_{\mathcal Q, \ell}^w \\
	& = \frac1c \sum_{\ell = 0}^{c-1}  1 - (1 - \FP_{\mathcal F, c})^{L-1}(1 - \FP_{\mathcal F}, \ell) \\
	& = 1 -  (1 - \FP_{\mathcal F, c})^{L-1}\sum_{\ell = 0}^{c-1} (1-\FP_{\mathcal F, \ell})
\end{align*}

\end{proof}

@@

	with the $u_\eta= p_{\eta}\FN_{\mathcal F,\eta} + \left(1-p_{\eta}\right)\left(1-\FP_{\mathcal F,\eta}\right)$.
\end{theorem}

\begin{proof}
\begin{align*}
	\FNR_{\mathcal Q, cn}^w 
	& = \frac{1}{cn} \sum_{k=1}^{cn} \FN_{\mathcal Q, k}^w 
	 = \frac{1}{cn} \sum_{k=1}^{n} \sum_{\ell = 0}^{c-1} \FN_{\mathcal Q, k + \ell}^w 
	= \frac1c \sum_{\ell = 0}^{c-1} \FN_{\mathcal Q, \ell}^w \\
	& = \frac1c \sum_{\ell = 0}^{c-1} u_c^{L-1} u_{\ell} 
	 = \frac{u_c^{L-1}}{c}  \sum_{\ell = 0}^{c-1}u_\ell
\end{align*}

\end{proof}

\section{Optimising queuing filters}
\todo[À vérifier je suis pas convaincu encore]
\begin{theorem}
	In first approximation, a low value of $L$ for constant value $cL$ leads to a higher false negative probability for the queueing filter.
\end{theorem}

@@

	On the other hand, a duplicate element $e^\star \dup E$ is classified as \UNSEEN if and only if it was not seen in the last $M$ elements of the stream, and the filter answers \UNSEEN. Let $D$ be the event \enquote{\emph{There is at least one duplicate in the stream}} and $C$ be the event \enquote{\emph{There is a duplicate of $e^\star$ in the $M$ previous elements of the stream}}. Then $e^\star$ triggers a false negative with probability
	\begin{align*}
	\FN_n &= (1 - \Pr[C | D])(1-p) =  \left(1 - \frac{\Pr[C \cap D]}{\Pr[D]}\right)(1-p)\\
	& = \left(1 - \frac{\Pr[C]}{\Pr[D]}\right)(1-p) =  \left(1 - \frac{1 - \Pr[\bar C]}{1 - \Pr[\bar D]}\right)(1-p) \\
	\FN_n &=  \left(1 - \frac{1 - \left(1 - \frac 1{|\Gamma|}\right)^M}{1 - \left(1 - \frac 1{|\Gamma|}\right)^n}\right)(1-p)
	\end{align*}
	Hence, the error probability of the filter is
	\begin{align*}
		ER_n 
		& = \FN_n + p \\
		& =  \left(1 - \frac{1 - \left(1 - \frac{1}{|\Gamma|}\right)^{M}}{1 - \left(1 - \frac 1{|\Gamma|}\right)^n}\right)(1-p) + p \\
		& = 1 - \frac{1-\left(1 - \frac{1}{|\Gamma|}\right)^{M}}{1 - \left(1 - \frac 1{|\Gamma|}\right)^n}(1-p),
	\end{align*}
	which is minimized when $p = 0$. 
	\end{proof}

Note, as highlighted in the proof, that this bound is \emph{not tight}: better bounds may exist, the study of which we leave as an open question for future work.

The results of an experimental comparison of different DDFs (details about the benchmark are given in Section~\ref{sub:saturation}) are summarizes in Figure~\ref{fig:graph_n}. It appears that the two most efficient filters (in terms of saturation rate) are the QHT \cite{10.1145/3297280.3297335} and the A2 \cite{Yoo10}.

@@

As noted in Section~\ref{sub:related}, A2 filters are a primitive form of a queuing filter (with $L=2$ and Bloom filters as subfilters) and thus were not included in this benchmark, as we focused on adapting filters designed for DDP to the sliding window DDP. This raises the question of whether recursive constructions (queuing filters of queuing filters) have particularly interesting properties, which we leave open for future work.

%In the experiments we use QHT for practical benchmarking of the queueing filter.
